---
title: "Homework 8: Basic Modeling Practice"
author: "Kim Costello"
format: html
editor: visual
---

### Seoul Bike Sharing Demand Dataset

```{r}
library(readr)
library(broom)
library(lubridate)
library(janitor)
library(dplyr)
library(ggplot2)
library(GGally)
```

The following data set is provided by the UCI Machine Learning Repository. The data set contains count of public bicycles rented per hours in the Seoul Bike Sharing System, with corresponding weather data and holiday information.

The following variables are provided:

-   Date - day/month/year

-   Rented Bike count - Count of bikes rented at each hour

-   Hour - Hour of the day

-   Temperature-Temperature in Celsius

-   Humidity - %

-   Windspeed - m/s

-   Visibility - 10m

-   Dew point temperature - Celsius

-   Solar radiation - MJ/m2

-   Rainfall - mm

-   Snowfall - cm

-   Seasons - Winter, Spring, Summer, Autumn

-   Holiday - Holiday/No holiday

-   Functional Day - NoFunc(Non Functional Hours), Fun(Functional hours)

```{r}
# Read in data
bike_data <- read_csv("SeoulBikeData.csv", locale = locale(encoding = "latin1"))
```
#### EDA

```{r}
# check for missing values
colSums(is.na(bike_data))
```

The data provided does not contain any missing values.

```{r}
str(bike_data)
```

```{r}
summary(bike_data)
```

```{r}
unique(bike_data$Seasons)
unique(bike_data$Holiday)
unique(bike_data$`Functioning Day`)
```

```{r}
bike_data <- bike_data |>
  mutate(
     Date = dmy(Date),
     Seasons = as.factor(Seasons),
     Holiday = as.factor(Holiday),
     `Functioning Day` = as.factor(`Functioning Day`)
  ) |>
  clean_names("snake")
  
```

```{r}
# Two-way contingency tables
table(bike_data$functioning_day, bike_data$seasons)
table(bike_data$functioning_day, bike_data$holiday)
```

```{r}
bike_data |>
  group_by(functioning_day) |>
  summarize(across(where(is.numeric), 
                   list("mean" = mean, "median" = median), 
                   .names = "{.fn}_{.col}"))
```

```{r}
bike_data |>
  group_by(functioning_day) |>
  summarize(min_bike_count = min(rented_bike_count), max_bike_count = max(rented_bike_count))
  
```

No bikes were rented during non-functional hours.

```{r}
fun_bike_data <- bike_data |>
  filter(functioning_day == "Yes")
```

```{r}
bike_summary_data <- fun_bike_data |>
  group_by(date, seasons, holiday) |>
  summarize(
    sum_bike_count = sum(rented_bike_count),
    sum_rainfall = sum(rainfall_mm),
    sum_snowfall = sum(snowfall_cm),
    across(c(temperature_c, 
             humidity_percent,
             wind_speed_m_s,
             visibility_10m,
             dew_point_temperature_c,
             solar_radiation_mj_m2), 
           ~ mean(.x),
           .names = "mean_{.col}")
    
  )

head(bike_summary_data)

    
```

```{r}
table(bike_summary_data$seasons, bike_summary_data$holiday)
```

```{r}
bike_summary_data |>
  group_by(seasons) |>
  summarize(across(where(is.numeric), 
                   list("mean" = mean, "median" = median), 
                   .names = "{.fn}_{.col}"))
```

```{r}
bike_summary_data |>
  group_by(seasons) |>
  summarize(min_bike_count = min(sum_bike_count), max_bike_count = max(sum_bike_count))
```

```{r}
df_corr <- bike_summary_data |>
  select(where(is.numeric)) |>
  rename("Bike_count" = sum_bike_count,
         "Rainfall" = sum_rainfall,
         "Snowfall" = sum_snowfall,
         "Temperature" = mean_temperature_c,
         "Humidity" = mean_humidity_percent,
         "Windspeed" = mean_wind_speed_m_s,
         "Visibility" = mean_visibility_10m,
         "Dew_point" = mean_dew_point_temperature_c,
         "Solar_radiation" = mean_solar_radiation_mj_m2)

ggcorr(df_corr, 
       nbreaks = 6,
       label = TRUE,
       palette = "BuPu",
       size = 2,          # adjusting variable names sizes
       hjust = 0.65) +    # adjusting position of variable names) 
  labs(title = "Correlation Matrix of all Numeric Variables") +
  ggeasy::easy_center_title()
```

```{r}
df_corr |>
  group_by(seasons) |>
  summarize(correlation = cor(Bike_count, Temperature, use = "pairwise.complete.obs"))
```

```{r}
g <- ggplot(data = bike_summary_data)

g + geom_point(aes(x = sum_bike_count, 
                   y = sum_rainfall, 
                   color = seasons)) +
  labs(x = "Rental Bike Count", 
       y = "Rainfall (mm)",
       title = "Rental Bike Count by Total Daily Rainfall",
       color = "Seasons") +
  scale_color_manual(values = c("#FFFF00",
                                "#FF00FF",
                                "#7FFF00",
                                "#00FFFF")) +
  theme_dark() +
  ggeasy::easy_center_title() 
```

```{r}
g + geom_boxplot(aes(x = seasons, y = sum_bike_count, fill = holiday)) +
  labs(x = "Seasons", 
       y = "Daily Rental Bike Count",
       title = "Daily Rental Bike Count by Seasons",
       fill = "Holiday") +
  ggeasy::easy_center_title() +
  scale_fill_manual(values = c("lightpink",
                               "lightblue"))
```

```{r}
ggplot(bike_summary_data, aes(x = date, y = sum_bike_count)) +
  geom_point() +
  labs(x = "Date", 
       y = "Daily Rental Bike Count",
       title = "Daily Rental Bike Count by Date") +
  ggeasy::easy_center_title() +
  geom_smooth()
```

#### Splitting the Data
